\section{Introduction}

Generally, this seminar (series) is about ``software languages'' (SL)
and hence it combines elements of software engineering, modelling,
programming languages, linguistics, formal specification, and
programming theory; see a recent discussion of some relevant fields
and terminology~\cite{FavreGLP10}.

Each particular edition of the SL seminar focuses on a specific topic
within the broader area of SL. (For instance, an edition may focus on
programming-language usage \cite{HageK08} or natural language aspects
in programming \cite{HostO08}.) The present notes provide some general
advice on the reporting and presentation obligations of the seminar's
participants. These notes are meant to be generic with regard to the
particular theme of a specific seminar edition.

That is, the purpose of these notes is to help student participants
writing their reports and preparing their presentations. These notes
complement the meetings and consultations.

\subsection*{Road-map}

In \S\ref{S:report}, advice on the report part of the seminar is
provided.

\noindent
In \S\ref{S:presentation}, advice on the presentation part
is provided.

\noindent
These notes are summarized in \S\ref{S:concl}.

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Advice on the report}
\label{S:report}

Much has been written and said about scientific writing.  A report for
this seminar is a small exercise in scientific writing. Hence,
seminar participants are \emph{strongly encouraged} to gain basic
understanding of scientific writing. For instance, seminar
participants may consult a resource such as
\cite{Zobel09,Day98}. Unfortunately, the syllabus of Koblenz seminars
does not plan for education on scientific writing, neither does any
other obligatory course. Hence, seminar participants are bound to
consult a designated book or online resource, and to work closely
together with the prof to get a reasonable result.\footnote{Here is
  another resource on writing and giving talks:
  \url{http://research.microsoft.com/en-us/um/people/simonpj/papers/giving-a-talk/giving-a-talk.htm}.}
The following notes try to get seminar participants off the ground.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Organization of the report}

As far as reports for the seminar are concerned, we assume a rather
simple organization. A report would be spending about 10-25 pages in
Springer's LNCS style to describe the research work carried out.

Reports should use close 10 pages if there are not any substantial
illustrations. Reports can use up to 25 pages, but \textbf{a higher
  page count does not imply a better grade}---potentially it may
negatively affect the grade. However, illustrations are strongly
encouraged, if they truly illustrate the subject at hand. So seminar
participants are welcome to exhaust a 25 pages limit with useful
illustrations included.

A report would have a title and an abstract. The purpose of an
abstract is to properly summarize the report: use 50-200 words,
typically 100-150 words.

A report consists of an introduction, several technical sections,
possibly a related work section, and a conclusion section. See below
for more details.

\subsection{The importance of the introduction}

A report should start with an \emph{introduction section} which
motivates the subject of the report, possibly states one or more
\emph{research questions}, typically provides some background material
(if such material is not placed in a subsequent, designated section or
an appendix), lists \emph{contributions} of the report, and describes
the organization of the rest of the report. Judgement of all submitted
work may be biased by the introduction.

If there is any ingredient that is critical to an excellent completion
of the seminar, then this is the formulation of a \emph{research
  question}. By default (for such a short report/paper that is meant
here), the research question should be identified in the introduction
section. The challenge is here to formulate a question that performs
well in all of the following dimensions: a) concise and clear; b)
informed and nontrivial; c) operationally intelligent and verifiable;
d) meaningful to the rest of the text.

\subsection{The technical sections}

After the introduction, there can be any number of \emph{technical
  sections}.  These sections constitute the core (``technical meat'')
of the report. The style of these technical sections very much depends
on the subject and the chosen sources as well as the assumed style of
presentation.

As there are many different styles, and we do not want to discuss them
here at length. The best advice may be to leverage the corpus of
consulted research papers and to try approximating the style of one of
them. The prof wants to help on this matter---during the consultations
and otherwise.

\subsection{The concluding section}

After the technical sections, the report should be concluded by a
final section. Typically, such a section contains bits of summary,
discussion, interpretations, open problems or future work, and
vision. This may be a good place for the student to show that he or
she has outgrown the view of one or two specific papers and can
usefully operate in a broader research context. Again, an excellent
student would make a serious effort to connect several research papers
and research directions in a comprehensible, informative, and
cross-indexing manner.

\subsection{Literature references}

One important element of scholarly writing is to make appropriate use
of citations. For instance, in the introduction section, one needs to
provide \emph{contextual citations} so that the overall context of the
research effort at hand is clarified. One may also quickly mention
some more technical sources as a means to prepare for the more
detailed parts of the report. Throughout the paper, one uses
\emph{technical citations} for two major purposes: (i) to give credit
to any reused content; (ii) to substantiate claims. In the conclusion
of the report, one uses \emph{visionary citations} to connect to
subjects that were not but cut be covered, or that are worth being
mentioned otherwise.

Seminar participants must prove that they use advanced means of
locating prior work. The \href{http://portal.acm.org/}{ACM Digital
  Library} and the
\href{http://www.informatik.uni-trier.de/~ley/db/}{DBLP} service are
particularly suited to search for scientific publications on topics of
the SL seminar.

Every edition of the SL seminar is unique, but here are some general
pointers for research on software languages; again, every edition of
the SL seminar may draw from additional sources, as it should be
evident from the seminar edition's design. Thus:

\begin{itemize}
\item SLE---Software Language Engineering
\item ASE---Automated Software Engineering
\item ICPC---International Conference on Program Comprehension
\item POPL---Principles of Programming Languages
\item OOPSLA---OO Programming, Systems, Languages, and Applications
\item MODELS--- Model Driven Engineering Languages and Systems
\item WCRE---Working Conference on Reverse Engineering
\item ICSM---International Conference on Software Maintenance
\item MSR---Mining Software Repositories
\item PLDI---Programm Language Design and Implementation
\item PPDP---Principles and Practice of Declarative Programming
\item VL/HCC---Symposium on Visual Languages and Human-Centric Computing
\end{itemize}

There are also various journals on similar themes that may provide
seminal and archival content for the seminar work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Abstraction is king, not page count}

There is no use in reproducing or condensing one specific paper (or a
series of strongly connected papers). Any reasonable seminar report
must exercise abstraction; it must study a subject from an individual
point of view with an original research question in mind. A reasonable
seminar report involves (3+) different pieces of scholarly work and
integrates them in the presentation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Advice on the presentation}
\label{S:presentation}

This section provides another example of what should be called a
technical section. Again, in such sections, the author works out the
technical details (say, meat) of all research work. There can be any
number of such sections. One should think hard about organization
principles to make up the appropriate number of sections and
subsections.

This section briefly talks about the presentation part of the seminar.
Just as in the case of scientific writing, there exists online or
published advice on ``how to give a good talk'', perhaps even in
seminar-like circumstances \cite{Talk,Klaeren94}. Giving good talks is
extremely difficult, but also very rewarding. Speakers should sure to
do some amount of reflection before they go into your talk. Some rules
of thumb follow:

\begin{itemize}

\item Finalizing the slides the night before is perhaps Ok, but one
  should start thinking of the talk way ahead of time. Otherwise, the
  presentation may likely to be boring or stressful and sends the
  audience to sleep. Speakers should not waste the precious cycles of
  the audience! As an aside, the available time should be perfectly
  leveraged by the talk: not more, not less.

\item One should not confuse writing with talking. One should not
  think of the talk as a way to present the paper in a condensed
  style. Instead, one should design the talk in a way that its
  structure and style maintains the attention of the audience and
  keeps up with the audience's abilities. One should use the talk to
  get people interested in the subject. One should motivate the
  audience to read more about the subject in the report that is
  advertised by the presentation.

\item One should not waste time on outlines; they are often totally
  useless at the time they are shown. One should focus on motivation
  from the first slide on. One should make good use of illustrations
  and simplify matters. One should not try to tell every detail, but
  to select some interesting details and make clear what else could
  be discussed or is included into the paper.

\item Seminar talks should be interactive. Hence, clarifying
  questions should be admitted.

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Concluding remarks}
\label{S:concl}

These notes are supposed to be useful as is, but further efforts may
be needed to convey all witness and intentions. For the time being,
the author asks the seminar participants and other readers to work
with what is available and the quoted resources. Everyone is welcome
to suggest revisions for these notes.
